API design for machine learning software: experiences from the scikit-learn project

### 2013-06-28 paper.tex
copyedit Comparison section

Other existing machine learning packages
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process input/output,
and will do so in a variety of languages.
The focus of \sklearn to incorporate the entire process into a single package still
allows for the use of external tools to process input/output, however gives the added
benefit of being able to run these processes within the same development environment.
In addition, this makes it easier for researchers and developers
to exchange and collaborate on software, as dependencies and setup are kept to a
minimum.

In comparison to specialized languages for numeric and statistical programming
such as \textsc{matlab} and R \citep{r},
Python has the distinct advantage
that it is a \textit{general purpose} language,
while NumPy and SciPy extend it with functionality
similar to that offered by \textsc{matlab} and R.
Python has strong language and standard library support for such tasks as
string/text processing, interprocess communication, networking
and many of the other auxiliary tasks that machine learning programs
(whether academic or commercial) routinely need to perform.
While support for many of these tasks is improving in languages such as
\textsc{matlab} and R, they still lag behind Python in their general purpose
applicability.
In many applications of machine learning these tasks, such as data access,
data preprocessing and reporting, can be a more significant task than applying
the actual learning algorithm.

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim},
which exemplifies API design for scalable processing of ``big data''.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorizers part of \sklearn
also use iterables to some extent,
they still produce entire sparse matrices, intended to be used for batch or
mini-batch learning. This is the case
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and allows effective use of the vectorized operations provided by NumPy and
SciPy.  We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.

>>>>

Other existing machine learning packages
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process input/output,
and will do so in a variety of languages.
By contrast, \sklearn allows users to implement that entire workflow
in a single program, written in a single language,
and developed in a single working environment.
This also makes it easier for researchers and developers
to exchange and collaborate on software, as dependencies and setup are kept to a
minimum.

Similar benefits hold in the case of specialized languages
for numeric and statistical programming
such as \textsc{matlab} and R \citep{r}.
In comparison to these, though, Python has the distinct advantage
that it is a \textit{general purpose} language,
while NumPy and SciPy extend it with functionality
similar to that offered by \textsc{matlab} and R.
Python has strong language and standard library support for such tasks as
string/text processing, interprocess communication, networking
and many of the other auxiliary tasks that machine learning programs
(whether academic or commercial) routinely need to perform.
While support for many of these tasks is improving in languages such as
\textsc{matlab} and R, they still lag behind Python in their general purpose
applicability.
In many applications of machine learning these tasks, such as data access,
data preprocessing and reporting, can be a more significant task than applying
the actual learning algorithm.

Within the realm of Python,
a package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim},
which exemplifies a different style of API design
geared toward scalable processing of ``big data''.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorizers part of \sklearn
also use iterables to some extent,
they still produce entire sparse matrices, intended to be used for batch or
mini-batch learning. This is the case
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and allows effective use of the vectorized operations provided by NumPy and
SciPy.  We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.
